Notes: New gaugings in 3d from three-form

### 2022-06-24 Notes_3dH3.tex
full IIB and beginning of 6d

\section{Truncations within \texorpdfstring{$\SL(8)_{\rm IIB}\subset\E_{8(8)}$}{SL(8)IIB in E8(8)}}

@@

\paragraph{\boldmath $H_{(7)}^{\alpha}$ and $H_{(5)}$ in $\SL(8)_{\rm IIB}$} Now that we know what are the $\SL(8)_{\rm IIB}\times\R_{+}$ representation corresponding to the internal coordinates, $C_{(6)}^{\alpha}$ and $C_{(4)}$, we can infer those of $H_{(7)}^{\alpha}$ and $H_{(5)}$:
\begin{equation}
  \begin{aligned}
    H_{(7)}^{\alpha} & \subset \bm{\bar{28}}_{2} \otimes \left(\bm{8}_{3}\oplus\bm{63}_{0}\right) = \bm{\bar{8}}_{5} \oplus \bm{\bar{216}}_{5} \oplus \bm{\bar{28}}_{2} \oplus \bm{\bar{36}}_{2} \oplus \bm{\bar{420}}_{2} \oplus \bm{\bar{1280}}_{2}, \\
    H_{(5)} & \subset \bm{\bar{28}}_{2} \otimes \bm{56}_{1} = \bm{8}_{3} \oplus \bm{216}_{3} \oplus \bm{1344}_{3}.
  \end{aligned}
\end{equation}

\paragraph{\boldmath $H_{(7)}^{\alpha}$ and $H_{(5)}$ in the embedding tensor}

\ce{Give decomposition of the embedding tensor to identify components, and count singlets.}

>>>>

\section{\texorpdfstring{${\rm AdS}_{3}\times S^{3}$}{AdS3xS3}: a proof of concept}
We start by reproducing the analysis of ref.~\cite{Eloy:2021fhc} to prove the validity of our method. We consider reductions within the minimal ${\cal N}=(1,0)$ six-dimensional supergravity, leading to duality group $\SO(4,4)$ in three dimensions, from which the half-maximal cases ${\cal N}=(1,1)$ and ${\cal N}=(2,0)$ can be deduced (in those cases the duality group in three dimensions in $\SO(8,4)$).

The coordinates $X^{[MN]}$ of the $\SO(p,q)$ exceptional field theory of ref.~\cite{Hohm:2017wtr} sit in the adjoint $\bm{28}$ of $\SO(4,4)$. Under $\SO(4,4)\rightarrow\SO(3,3)\times\R_{+}$, it decomposes into
\begin{equation}
  \begin{aligned}
    \bm{28} &\longrightarrow \bm{6}_{2} \oplus \bm{1}_{0} \oplus \bm{15}_{0} \oplus \bm{6}_{-2},\\
    X^{MN} &\longrightarrow \{Y^{A0},Y^{0}{}_{0},Y^{[AB]},Y^{A}{}_{0}\}.
  \end{aligned}
\end{equation}
Further decomposing $\SO(4,4)\rightarrow\SO(3,3)\times\R_{+}\rightarrow\SL(3)\times\R_{+}\times\R_{+}$, we get
\begin{equation}
  \bm{28} \longrightarrow \left[\bm{\bar{3}}_{2,2} \oplus \bm{3}_{-2,2}\right] \oplus \bm{1}_{0,0} \oplus \left[\bm{3}_{4,0} \oplus \bm{1}_{0,0} \oplus \bm{8}_{0,0} \oplus \bm{\bar{3}}_{-4,0}\right] \oplus \left[\bm{\bar{3}}_{2,-2} \oplus \bm{3}_{-2,-2}\right].
\end{equation}
They are two physically different possibilities to embed the internal coordinates in $X^{MN}$, corresponding to the two different half-maximal supergravities in six dimensions:\footnote{The choices are equivalent in $\SO(4,4)$ (through triality), but different once embedded in $\SO(8,4)$.}
\begin{equation}
  \begin{aligned}
    {\cal N}=(1,1):&\qquad y^{m}=Y^{m0}\quad {\rm in}\quad  \bm{\bar{3}}_{2,2} \subset \bm{6}_{2},\\
    {\cal N}=(2,0):&\qquad \tilde{y}_{m}=\varepsilon_{mnp}Y^{np}\quad {\rm in}\quad  \bm{3}_{4,0} \subset \bm{15}_{0}.
  \end{aligned}
\end{equation}

\section{Truncations within \texorpdfstring{$\SL(8)_{\rm IIB}\subset\E_{8(8)}$}{SL(8)IIB in E8(8)}}
We analyse here whether the type IIB $S^{7}$ truncation of ref.~\cite{Galli:2022idq} admit additional gaugings.

@@

\paragraph{\boldmath $H_{(7)}^{\alpha}$ and $H_{(5)}$ in $\SL(8)_{\rm IIB}$} Now that we know what are the $\SL(8)_{\rm IIB}\times\R_{+}$ representations corresponding to the internal coordinates, $C_{(6)}^{\alpha}$ and $C_{(4)}$, we can infer those of $H_{(7)}^{\alpha}$ and $H_{(5)}$:
\begin{equation}  \label{eq:H7H5inIIB}
  \begin{aligned}
    H_{(7)}^{\alpha} & \subset \bm{\bar{28}}_{2} \otimes \left(\bm{8}_{3}\oplus\bm{63}_{0}\right) = \bm{\bar{8}}_{5} \oplus \bm{\bar{216}}_{5} \oplus \bm{\bar{28}}_{2} \oplus \bm{\bar{36}}_{2} \oplus \bm{\bar{420}}_{2} \oplus \bm{\bar{1280}}_{2}, \\
    H_{(5)} & \subset \bm{\bar{28}}_{2} \otimes \bm{56}_{1} = \bm{8}_{3} \oplus \bm{216}_{3} \oplus \bm{1344}_{3}.
  \end{aligned}
\end{equation}

\paragraph{\boldmath $H_{(7)}^{\alpha}$ and $H_{(5)}$ in the embedding tensor} Under $\E_{8(8)}\rightarrow\SL(9)\rightarrow\SL(8)_{\rm IIB}\times\R_{+}$, the $\E_{8(8)}$ embedding tensor decomposes as
\begin{equation}
  \begin{aligned}
    \bm{3875} &\longrightarrow {\color{mydarkorange}\bm{80}} \oplus \bm{240} \oplus {\color{mydarkorange}\bm{\bar{240}}} \oplus \bm{1050} \oplus {\color{mydarkorange}\bm{\bar{1050}}} \oplus {\color{mydarkorange}\bm{1215}} \\
    &\longrightarrow \left[{\color{mydarkorange}\bm{8}_{3}} \oplus \bm{1}_{0} \oplus \bm{63}_{0} \oplus \bm{\bar{8}}_{-3}\right] \oplus \left[\bm{168}_{1} \oplus \bm{28}_{-2} \oplus \bm{36}_{-2} \oplus \bm{8}_{-5}\right] \oplus \left[{\color{mydarkorange}\bm{\bar{8}}_{5}} \oplus {\color{mydarkorange}\bm{\bar{28}}_{2}} \oplus {\color{mydarkorange}\bm{\bar{36}}_{2}} \oplus \bm{\bar{168}}_{-1} \right] \\
    &\quad \oplus \left[\bm{70}_{4} \oplus \bm{56}_{1} \oplus \bm{504}_{1} \oplus \bm{420}_{-2}\right] \oplus \left[{\color{mydarkorange}\bm{\bar{420}}_{2}} \oplus \bm{\bar{56}}_{-1} \oplus \bm{\bar{504}}_{-2}\oplus \bm{\bar{70}}_{-4} \right] \oplus \left[{\color{mydarkorange}\bm{216}_{3}} \oplus \bm{63}_{0} \oplus \bm{720}_{0} \oplus \bm{\bar{216}}_{-3}\right].
  \end{aligned}
\end{equation}
The coloured representations are those that can originate from $H_{(7)}^{\alpha}$ and $H_{(5)}$, \textit{i.e.} those in common with eq.~\eqref{eq:H7H5inIIB}. Among them, only the $\bm{\bar{36}}_{2}$ features a singlet under $\SO(8)_{\rm IIB}$. This is the singlet used in ref.~\cite{Galli:2022idq} to support the $S^{7}$ truncation, which thus do not admit further gaugings.
